---
title: "energy-data"
format: html
editor: visual
---

## Libraries

Here we install all the packages needed on local, if they are not installed before.

```{r}
packages_to_install <- c("tidytuesdayR", "dplyr", "tidyr", "ggmap", "rworldmap", "plotly", "readr")

for (package in packages_to_install) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)
  }
}

```

```{r}
library("tidytuesdayR")
library("dplyr")
library("tidyr")
library("ggmap")
library("rworldmap")
library("plotly")
library("readr")
```

## Loading Data

```{r}
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

# tuesdata <- tidytuesdayR::tt_load('2023-06-06')
tuesdata <- tidytuesdayR::tt_load(2023, week = 23)

owid_energy <- tuesdata$`owid-energy`

# Or rad in the data manually

# owid_energy <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-06/owid-energy.csv')


## Data on my local
# url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-06/owid-energy.csv"
# 
# loc <- "data/owid-energy.csv"
# 
# download.file(url, destfile = loc)
# 
# owid_energy <- read_csv(loc)
```

## `Exploring Data`

### `Column names`

```{r}
colnames(owid_energy)
```

## Maps

it is to zoom maps for different energies and maybe map them in different years.

```{r}
nuclear_data <- owid_energy |> 
  select(country, contains("nuclear"))
```

```{r}
# joinCountryData2Map(nuclear_data, joinCode="NAME", nameJoinColumn="country") |> 
# mapCountryData(nameColumnToPlot = "nuclear_consumption", mapTitle = "Nuclear Consumption by Country", colourPalette="heat", oceanCol = "#CCCCCCCC", addLegend = TRUE, aspect = 1.1, borderCol = "Black", lwd =.1)
# # terawatt-hours
```

## Greenhouse gas emissions vs GDP

GDP: Total real gross domestic product, inflation-adjusted

It is the standard measure of the value added created through the production of goods and services in a country during a certain period. As such, it also measures the income earned from that production, or the total amount spent on final goods and services (less imports)

we need #gdp per capita for later

```{r}
gge_data <-
  owid_energy |>
    select(country, year, gdp, greenhouse_gas_emissions) |>
    drop_na()

gge_data |> 
  summarise_all(n_distinct)
```

```{r}
gge_data |> 
  arrange(desc(gdp))
```

```{r}
df <- 
  gge_data |> 
    filter(country != "World")
df
```

```{r}
df |> 
  filter(country == "China") |> 
  ggplot() + 
  aes(x= gdp, y = greenhouse_gas_emissions, color = year) +
  geom_point() + 
  labs(title = "greenhouse_gas_emissions vs gdp China")
```

```{r}
df |> 
  filter(country == "United States") |> 
  ggplot() + 
  aes(x= gdp, y = greenhouse_gas_emissions, color = year) +
  geom_point() + 
  labs(title = "greenhouse_gas_emissions vs gdp U.S.")

```

## Electricity Generation by Country

```{r}
fig <- plot_ly(
  data = owid_energy,
  x = ~country,
  y = ~electricity_generation,
  type = "bar",
  marker = list(color = "blue")
)

fig <- fig |>  layout(title = "Electricity Generation by Country")

fig

```

### ideas

with this plot I realized that in the country column we have individual countries (like Afghanistan, Albania, Algeria, etc.), regions (such as Africa, Asia, Europe), specific groupings or categories (like OECD, OPEC, High-income countries), and even historical or special entities (such as USSR, East Germany).

so one idea is to compare between aforementioned groups, the other idea is to analyse different types of energy such as:

```         
biofuel
coal
electricity
Primary energy
fossil fuel
gas 
hydropower 
low-carbon energy
nuclear power
oil
renewable
solar power
wind power
```

There is a codebook explaining every column:

<https://github.com/owid/energy-data/blob/master/owid-energy-codebook.csv>

lots of different fun analyses can be made!

## Solar Energy Consumption Trends

```{r}
Selected_Countries <- c("Germany", "United States", "China")

filtered_energy_data <- owid_energy |>
  filter(country %in% Selected_Countries) |>
  select(country, year, solar_consumption) |>
  pivot_longer(cols = solar_consumption, 
               names_to = "energy_type", values_to = "consumption")

plot <- filtered_energy_data |>
  plot_ly(x = ~year, y = ~consumption, color = ~country, 
          type = 'scatter', mode = 'lines+markers',
          text = ~paste("Country:", country, "<br>", "Year:", year, "<br>", "Solar Consumption:", consumption),
          hoverinfo = "text") |>
  layout(title = "Solar Energy Consumption Trends",
         xaxis = list(title = "Year", tickangle = 90),
         yaxis = list(title = "Solar Energy Consumption (TWh)"),
         colorway = RColorBrewer::brewer.pal(3, "Dark2"))

plot
```

## Renewable Energy Consumption Trends

```{r}
Selected_Country <- "Germany"

filtered_energy_data <- owid_energy |>
  filter(country == Selected_Country) |>
  select(year, solar_consumption, wind_consumption, biofuel_consumption, other_renewable_consumption, hydro_consumption) |>
  pivot_longer(cols = c(solar_consumption, wind_consumption, biofuel_consumption, other_renewable_consumption, hydro_consumption), 
               names_to = "energy_type", values_to = "consumption")

plot <- filtered_energy_data |>
  plot_ly(x = ~year, y = ~consumption, color = ~energy_type, 
          type = 'scatter', mode = 'lines+markers',
          text = ~paste("Year:", year, "<br>", "Consumption:", consumption),
          hoverinfo = "text") |>
  layout(title = sprintf("Renewable Energy Consumption Trends in the %s", Selected_Country),
         xaxis = list(title = "Year", tickangle = 90, rangeselector = list(buttons = list(list(count = 1, label = "1y", step = "year", stepmode = "backward"), list(count = 5, label = "5y", step = "year", stepmode = "backward"), list(step = "all"))), rangeslider = list(type = "date")),
         yaxis = list(title = "Energy Consumption (TWh)"),
         colorway = RColorBrewer::brewer.pal(5, "Dark2"))

plot
```

```{r}
Selected_Countries <- c("Germany", "United States", "China")

filtered_energy_data <- owid_energy |>
  filter(country %in% Selected_Countries) |>
  select(country, year, solar_consumption, wind_consumption, biofuel_consumption, other_renewable_consumption, hydro_consumption) |>
  pivot_longer(cols = -c(country, year), names_to = "energy_type", values_to = "consumption")

ggplot(filtered_energy_data, aes(x = year, y = consumption, fill = energy_type)) +
  geom_area(position = 'stack') +
  facet_wrap(~ country) +
  labs(title = "Stacked Area Chart of Energy Consumption by Type",
       x = "Year",
       y = "Energy Consumption (TWh)",
       fill = "Energy Type") +
  theme_minimal()

```

```{r}
data_2020 <- owid_energy |> 
  filter(year == 2018, !is.na(gdp), !is.na(electricity_demand), !is.na(population), country != "World") |> 
  mutate(size = sqrt(population * 3e-06)) 

colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

fig <- plot_ly(data_2020, x = ~gdp, y = ~electricity_demand, type = 'scatter', mode = 'markers',
               marker = list(size = ~size, sizemode = 'diameter', line = list(width = 2, color = '#FFFFFF')),
               text = ~paste('Country:', country, '<br>Electricity Demand:', electricity_demand, '<br>GDP:', gdp,
                             '<br>Population:', population)) |> 
  layout(title = 'Electricity Demand vs GDP',
         xaxis = list(title = 'GDP', type = 'log'),
         yaxis = list(title = 'Electricity Demand (TWh)'),
         plot_bgcolor = 'rgb(243, 243, 243)')

fig

```

in above plot, what is needed to be done, is to just choose a fee countries and compare them. becasue nothing is clear from this plot.

### How much of the country's electricity comes from renewables?

```{r}
filtered_data <- owid_energy |> 
  select(year, matches("_electricity$"), -other_renewable_exc_biofuel_electricity, -other_renewable_electricity, -per_capita_electricity)
colnames(filtered_data)
```

```{r}

filtered_data <- owid_energy |> 
  filter(country == "United States") |> 
  select(year, matches("_electricity$"), -other_renewable_exc_biofuel_electricity, -other_renewable_electricity, -per_capita_electricity)

long_data <- filtered_data |> 
  pivot_longer(cols = -year, names_to = "electricity_type", values_to = "production")

fig <- plot_ly(data = long_data, x = ~year, y = ~production, color = ~electricity_type, type = 'scatter', mode = 'lines+markers') %>%
  layout(title = "Electricity Production Over Time by Type",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Electricity Production (TWh)"))

fig



```

## Groups

```{r}
 unique_countries <- unique(owid_energy$country) 
 unique_countries
```

### OECD

In dataset, the entries "Non-OECD (BP)" and "Non-OECD (EIA)" refer to groups of countries that are not part of the Organisation for Economic Co-operation and Development (OECD), as defined by two different organizations: BP and the EIA.

```{r}
owid_energy |> 
  filter(country %in% c("OECD (EIA)", "Non-OECD (EIA)")) |> 
  ggplot(aes(x = year, y = energy_cons_change_pct, fill = country)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_minimal()

owid_energy |>
     filter(country %in% c("OECD (BP)", "Non-OECD (BP)") & year == "2024")
```

### Income Groups

in the owid-energy dataset, in country column we have different income groups such as:

```         
    "L": "Low-income countries"
    "H": "High-income countries"
    "UM": "Upper-middle-income countries"
    "LM": "Lower-middle-income countries"
```

source of pyhton code that did this grouping for owid data: <https://github.com/owid/etl/blob/master/etl/steps/data/garden/wb/2023-04-30/income_groups.py>

```{r}
owid_energy |> 
   filter(country %in% c("Upper-middle-income countries", 
                        "Low-income countries", 
                        "High-income countries", 
                        "Lower-middle-income countries"))
```

**Correlation Heatmaps**: Explore the relationships between different types of energy consumption, GDP, and population.
